---
title: "Compare covariate formulas with paired and nested CV"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compare covariate formulas with paired and nested CV}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(IntegMultiReg)
```

Clinical covariates are forced within an IMR specification; the selection prior
acts on molecular features. This example compares `y ~ age` with
`y ~ age + sex + stage` on synthetic continuous outcomes. Both candidates use
the current symmetric-MRF fitter and training-fold refit CV. The example demonstrates
predictive comparison, not causal confounder selection. Required adjustment
variables must appear in every candidate.

The [getting started guide](IntegMultiReg.html) shows the formula calls and
explicitly passes the first CV result's saved folds to the second candidate.
Here the installed example extends that procedure to nested selection. Its
single implementation generates the results below and the CSV audit files.

# Run the complete example

```{r run-comparison, results = "hide"}
source(system.file("examples", "compare-covariates.R", package = "IntegMultiReg"))
comparison <- run_covariate_comparison(
  out_dir = file.path(tempdir(), "covariate-comparison")
)
```

This retains 2,000 draws after 500 burn-in iterations in each fit. The paired
comparison uses two rounds of three-fold CV; the nested procedure uses three
outer folds and three inner folds. `quick = TRUE` reduces these settings for
a smoke test. Short runs or reproducible output do not establish chain
convergence or validate the illustrative covariates scientifically.

Earlier results from historical updates are kept with their source snapshots.
They are not silently relabelled as results from this workflow.

# Compare prespecified candidates

Both candidates receive exactly the same subject/fold assignments. Each fit
rebuilds preprocessing, formula encoding, selection MCMC and prediction inside
the training fold. The pooled MSE combines the held-out predictions for a round.

```{r paired-results}
comparison$paired_summary
comparison$paired_rounds
```

`expanded_minus_age_only` is the paired difference for each round; negative
values favor the expanded formula. The SD across these overlapping CV rounds
is descriptive and is not an independent-replicate standard error.

```{r paired-folds}
head(comparison$paired_folds)
stopifnot(identical(
  comparison$paired_cv_controls[[1]]$folds,
  comparison$paired_cv_controls[[2]]$folds
))
comparison$paired_cv_controls[[1]]$refit_seeds
```

`compare_fit_summaries()` supplies structural metadata. The predictive comparison above
comes from CV, not from a ranking performed by `compare_fit_summaries()`.

# Evaluate formula selection with outer folds

For each outer split, the script fits both candidates using only the outer
training subjects. Inner CV selects the smaller pooled MSE, with candidate
order breaking exact ties. The chosen fit then predicts the untouched outer
test subjects. The outer outcomes enter only the final scoring.

```{r nested-results}
comparison$selected_by_fold
comparison$nested_summary
```

The pooled outer MSE estimates the performance of this selection procedure
under its settings. It does not estimate the performance of the candidate
that happens to look best after inspecting all outer results.

The saved records let readers check subject separation and reconstruct the
score without rerunning the sampler:

```{r audit-separation}
p <- comparison$outer_predictions
for (fold in unique(p$outer_fold)) {
  inner_ids <- comparison$nested_inner_folds$id[
    comparison$nested_inner_folds$outer_fold == fold
  ]
  stopifnot(setequal(inner_ids, p$id[p$outer_fold != fold]),
            !any(inner_ids %in% p$id[p$outer_fold == fold]))
}
stopifnot(isTRUE(all.equal(
  mean((p$observed - p$prediction)^2),
  comparison$nested_summary$pooled_outer_mse
)))
```

# Keep the record needed for replay

The output directory contains eight CSV tables, `comparison.rds` and
`sessionInfo.txt`. The RDS file also records the sampler convention, MCMC
settings, fit and outer-split seeds, and the complete CV controls for the paired
and nested fits, including actual folds and refit seeds. Keep the exact package
source and this record together. A matching version label or seed alone does
not identify the source or guarantee matching arithmetic across platforms.
