---
title: "Integrative Bayesian Multi-Platform Regression with IntegMultiReg"
author: "Sinian Zhang, Jianfeng Wang, Thierry Chekouo"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Integrative Bayesian Multi-Platform Regression with IntegMultiReg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.2
)
set.seed(1)
```

# Overview

`IntegMultiReg` implements the *integrative multi-regression* (IMR) model of
Chekouo, Stingo, Doecke and Do (2017) and extends it from time-to-event
outcomes to continuous (Gaussian) and binary (probit) outcomes.  The central
idea is to integrate **all** available subjects across several molecular
platforms, even when many subjects are measured on only a subset of the
platforms.

Given $L$ platforms, subjects are partitioned into up to $2^L-1$ non-empty
**availability subgroups** of an $L$-set Venn diagram.  One regression model is
built per retained subgroup. Its active coefficients have **non-local
product-moment priors**, which downweight effects near zero. A **Markov random
field prior** on the variable-selection indicators shares information across
availability subgroups by encouraging common biomarkers to be selected.

```{r load}
library(IntegMultiReg)
data("simIMR", package = "IntegMultiReg")
sapply(simIMR$platforms, dim)
```

The bundled `simIMR` data set has three platforms (`genomic`, `proteomic`,
`metabolomic`) measured on overlapping but partially missing sets of subjects,
clinical covariates for everyone, and three outcome types generated from the
same latent signal.  The truly associated features are recorded in
`simIMR$truth`.

# Prepare and inspect a data object

The fitter uses the symmetric MRF update and its boundary Hastings correction.
The model variant is `"imr"` for MRF borrowing or `"bms"` for independent
subgroups. Historical updates are replayed from their archived source, rather
than selected through the ordinary fitting interface. These examples use
training-fold refit CV with hyperparameters fixed in advance.

Construct an `imr_data` object once, inspect it, and pass it to fitting or
prediction. The constructor checks identifiers, outcome coding and alignment.
Standardization is estimated during fitting within each retained subgroup.

```{r prepare-data}
analysis <- imr_data(
  platforms = simIMR$platforms,
  outcome = simIMR$outcome.binary,
  covariates = simIMR$covariates,
  outcome_type = "binary"
)
analysis
summary(analysis)
stopifnot(validate_imr_data(analysis))
head(as.data.frame(analysis))
```

# Fit the model

The workhorse is `imr()`.  Here we fit a binary (probit) outcome with a short
MCMC run for illustration; in practice use longer chains.

```{r fit}
fit <- imr(
  analysis,
  nu = c(-4, -3, -4),
  draws = 1500, burnin = 500,
  min_subgroup_size = 30,
  seed = 1
)
fit
```

The four modelled availability subgroups correspond to the availability patterns
`011` (genomic + proteomic), `111` (all three), `101` (genomic + metabolomic)
and `100` (metabolomic only); the bitstring digits run from the first platform
(right) to the last (left).  `plot_subgroup_sizes()` shows how the sample splits
across them.

```{r plot-sizes, fig.height = 3.6}
plot_subgroup_sizes(fit)
```

# Predict with a data object

Prediction data use the same class and need no outcome column. Here we reuse
rows to demonstrate routing to the genomic/proteomic subgroup; these fitted-data
predictions are not an assessment of predictive accuracy. New subjects must
supply the clinical columns required by the fit. Binary predictions are event
probabilities.

```{r predict}
new_data <- imr_data(
  platforms = list(
    genomic = simIMR$platforms$genomic[1:20, ],
    proteomic = simIMR$platforms$proteomic[1:20, ]
  ),
  covariates = simIMR$covariates
)
stopifnot(validate_imr_data(new_data))
pred <- predict(fit, newdata = new_data)
head(pred[["subgroup:011"]])
```

# Use a formula and data frame

The formula describes the outcome and forced clinical covariates. Molecular
platforms are supplied separately and remain subject to biomarker selection.
The identifier aligns subjects and is not a predictor. This continuous example
includes `age`, `sex` and `stage`; formula encoding is stored for prediction.

```{r formula-fit}
clinical <- merge(simIMR$outcome.continuous, simIMR$covariates,
                  by = "id", sort = TRUE)
formula_fit <- imr(
  y ~ age + sex + stage, data = clinical, platforms = simIMR$platforms,
  outcome_type = "continuous", nu = c(-4, -3, -4),
  draws = 1500, burnin = 500, min_subgroup_size = 30, seed = 1
)
head(predict(formula_fit, newdata = new_data)[["subgroup:011"]])
```

# Compare formulas on the same folds

Prespecify candidates and evaluate them on the same subjects. Passing saved
folds explicitly makes the comparison paired; matching seed numbers alone is
not a check of subject/fold assignments. Every clinical term is forced within
its candidate. Necessary adjustment terms must remain in every candidate.

```{r paired-formulas}
age_fit <- imr(
  y ~ age, data = clinical, platforms = simIMR$platforms,
  outcome_type = "continuous", nu = c(-4, -3, -4),
  draws = 1500, burnin = 500, min_subgroup_size = 30, seed = 1
)
cv_age <- cv_imr(age_fit, k = 3, rounds = 1, cv_method = "refit")
cv_full <- cv_imr(formula_fit, k = 3, rounds = 1, cv_method = "refit",
                  folds = cv_age$control$folds)
keys <- c("round", "id", "subgroup", "fold")
stopifnot(identical(cv_age$predictions[keys], cv_full$predictions[keys]))
data.frame(
  formula = c("y ~ age", "y ~ age + sex + stage"),
  pooled_mse = c(cv_age$pooled[1, "all"], cv_full$pooled[1, "all"])
)
```

These short runs demonstrate the interface. Selecting a formula using these
scores and reporting the same scores as its performance introduces selection
optimism. The [covariate comparison guide](covariate-comparison.html) runs the
complete paired comparison and nested selection workflow, with outer test
subjects excluded from every inner fold.

# Inspecting selected biomarkers

`summary()` ranks, per platform, the features whose marginal posterior
inclusion probability (mPIP) exceeds a threshold.

```{r summary}
summary(fit, threshold = 0.5)
```

The posterior inclusion probabilities themselves are available through `inclusion_probabilities()`
(a list of subgroup $\times$ feature matrices) and can be visualised either as
per-platform heatmaps or as a single ranked bar chart of the most strongly
supported features across all platforms.

```{r plot-selection}
plot(fit, type = "selection", platform = 1)
```

```{r plot-top, fig.height = 4}
plot_top_features(fit, top = 8)
```

The estimated MRF interaction parameters, which measure how strongly each pair
of subgroups shares biomarkers, are shown with `type = "theta"`, and the
log-posterior trace with `type = "trace"`.

```{r plot-trace}
plot(fit, type = "trace")
```

# Coefficient and predictive uncertainty

`selection_summary(fit)` summarizes selection indicators and MRF interactions
already stored in the fitted object. Regression coefficients require a second
step, `sample_regression_posterior()`. The following small continuous example shows that
step and keeps the printed coefficient table short.

```{r posterior-fit}
assay <- data.frame(id = 1:40, marker = seq(-1, 1, length.out = 40))
response <- data.frame(id = assay$id,
                       y = 1 + assay$marker + sin(assay$id) / 3)
uncertainty_fit <- imr(
  list(assay = assay), response, outcome_type = "continuous",
  forced_prior_scale = 1, min_subgroup_size = 5,
  draws = 500, burnin = 250, seed = 1
)
post <- sample_regression_posterior(
  uncertainty_fit, output_draws = 1000, burnin = 1000, chains = 4,
  min_draws_per_model_chain = 1000, seed = 2
)
post
post$diagnostics
confint(post)[[1]]
```

`draws` controls the number of returned model-averaged samples, while
`min_draws_per_model_chain` is the minimum retained length of each conditional chain.
Inspect the model-specific diagnostics before interpreting intervals. These
diagnostics assess conditional coefficient/variance sampling, not the original
selection chain. Model weights still inherit the fitted Laplace approximation.

```{r posterior-predict}
new_assay <- list(assay = data.frame(id = 41:43, marker = c(-0.5, 0, 0.5)))
predict(post, new_assay, quantity = "conditional_mean")
predict(post, new_assay, quantity = "new_observation")
```

The first call summarizes uncertainty in the conditional response mean. The
second also generates future outcome variation. `summary(post)` and
`confint(post)` read the same stored posterior draws; they do not rerun MCMC.
For binary or censored fits, `sample_regression_posterior(..., latent = TRUE)` additionally
retains the augmented response for inspection with
`summary(post, parm = "latent")`. A continuous outcome has no latent response.

# Assessing predictive performance

`cv_imr()` defaults to `cv_method = "refit"`. The alternative `"reweight"`
reuses a full-data fit; `model_set = "all_draws"` keeps empirical multiplicities,
while `"top_unique"` retains ranked distinct states. Reweighting uses full-fit
preprocessing and latent-response summaries. Held-out outcomes enter its
importance correction, which alone is not evidence of an error.

This example explicitly uses `cv_method = "refit"` to refit each training fold, including
standardization, selection MCMC and model weights. It reports AUC for binary
outcomes, concordance for survival and MSE for continuous outcomes, overall
and within each retained availability subgroup. Formula transformations are
also rebuilt on training rows. Runtime scales with folds times rounds; use
short chains for a code-path check and increase them for final analysis.
To compare IMR with BMS, fit a separate `model_variant = "bms"` object. The original
hyperparameters are treated as fixed; tuning requires another validation layer.

```{r cv}
cv <- cv_imr(fit, k = 5, rounds = 3, cv_method = "refit")
cv$metric
round(colMeans(cv$pooled), 3)
```

# Other outcome types

The same interface handles continuous and right-censored outcomes; only
`outcome_type` and the `outcome` data frame change.

```{r other-types, eval = FALSE}
# continuous (Gaussian) outcome
fit_c <- imr(
  simIMR$platforms, simIMR$outcome.continuous, covariates = simIMR$covariates,
  outcome_type = "continuous", nu = c(-4, -3, -4),
  draws = 1500, burnin = 500, min_subgroup_size = 30, seed = 1)

# right-censored survival outcome
fit_s <- imr(
  simIMR$platforms, simIMR$outcome.survival, covariates = simIMR$covariates,
  outcome_type = "right.censored", nu = c(-4, -3, -4),
  draws = 1500, burnin = 500, min_subgroup_size = 30, seed = 1)
```

# Real TCGA-KIRC example

The package also includes `kircIMR`, a reduced real-data example derived from
public UCSC Xena TCGA kidney renal clear cell carcinoma (KIRC) sampleMap files,
not from controlled-access TCGA/GDC files.  It mirrors the case-study structure
of Chekouo et al. (2017): mRNA expression, miRNA expression and DNA methylation
measured on overlapping patient sets, clinical covariates, and a right-censored
survival outcome.  Patient IDs are package-internal labels such as `KIRC001`;
the package does not distribute TCGA barcodes, and users should not attempt
participant re-identification or linkage to external resources.

```{r load-kirc}
data("kircIMR", package = "IntegMultiReg")
sapply(kircIMR$platforms, dim)
kircIMR$model_subgroup_sizes
```

The original Biometrics analysis used a much larger screened panel
(776 mRNA, 91 miRNA and 729 methylation features) and long MCMC chains.  The
package object keeps a smaller Cox-screened panel (50, 30 and 50 features) so
that examples remain lightweight while retaining the same scientific data
structure.  A short survival fit follows the same call pattern:

```{r kirc-fit, eval = FALSE}
kirc_fit <- imr(
  x = kircIMR$platforms,
  outcome            = kircIMR$outcome.survival,
  covariates = kircIMR$covariates,
  outcome_type       = "right.censored",
  nu                 = c(-4, -3, -4),
  draws = 4000, burnin = 1000,
  min_subgroup_size              = 30,
  seed = 1
)
summary(kirc_fit)
plot_top_features(kirc_fit, top = 12)
```

# Record computational settings

Use `folds = result$control$folds` for partition replay. Reweighting also retains
`row_order`, which preserves summation order. `fold_rng = "continue"` can resume
a compatible fit's saved GSL stream for reweighting; generated refit folds use
R's generator. Seeds alone do not identify the actual partitions.

The fit records prior scales and Laplace optimization controls. The
[replay guide](reproducibility.html) connects the exact source, data, settings,
folds and unrounded outputs. The
[methods guide](https://h7nian.github.io/IntegMultiReg/method-coverage.html) explains model approximations and
historical differences. Successful execution does not establish convergence
or reproduce an entire historical study.

# References

Chekouo T, Stingo FC, Doecke JD, Do K-A (2017). "A Bayesian Integrative
Approach for Multi-Platform Genomic Data: A Kidney Cancer Case Study."
*Biometrics*, **73**(2), 615--624. <https://doi.org/10.1111/biom.12587>
