---
title: "Get started with IntegMultiReg"
author: "Sinian Zhang, Jianfeng Wang and Thierry Chekouo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with IntegMultiReg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(IntegMultiReg)
```

IntegMultiReg fits Bayesian regressions to molecular platforms measured on
partially overlapping subjects. It groups subjects by their available platforms,
selects molecular predictors and estimates their effects. The IMR model shares
selection information across groups; BMS fits the groups independently.
Version 0.3.0 samples selection indicators, regression coefficients, residual
variances and sharing parameters in one joint MCMC run.

# Prepare, check and fit data

Each platform has an `id` column and numeric features. Outcome and clinical
covariate tables also use `id`; rows are aligned by identifier, not position.
An entirely unavailable platform for a subject is allowed. Missing measurements
within an otherwise available platform are rejected.

```{r prepare}
data(simIMR)
analysis <- imr_data(
  platforms = simIMR$platforms,
  outcome = simIMR$outcome.continuous,
  covariates = simIMR$covariates,
  outcome_type = "continuous"
)
validate_imr_data(analysis)
head(as.data.frame(analysis))
```

Keep statistical priors separate from simulation settings:

```{r fit}
priors <- imr_priors(nu = c(-4, -3, -4))
sampling <- imr_mcmc(draws = 2000, burnin = 1000, chains = 4, seed = 123)
fit <- imr(analysis, priors = priors, mcmc = sampling)
fit
```

`draws` is the number retained **per chain**. The four chains here retain 8,000
joint states. These settings illustrate the workflow; inspect the diagnostics
before interpreting the posterior. A warning is evidence to reassess the run,
not a request to suppress diagnostic computation.

```{r inspect}
validate_imr_object(fit)
head(mcmc_diagnostics(fit))
lapply(coef(fit), head)
head(confint(fit))
```

`coef(fit)` returns posterior mean regression coefficients. `confint(fit)`
returns equal-tailed credible intervals from the same joint draws, including
exact zeros when a molecular feature is excluded. These are posterior
intervals for effects, not confidence intervals for estimated inclusion
probabilities.

For a feature j in subgroup s, the marginal inclusion probability is estimated by

\[
\widehat{P}(\gamma_{sj}=1\mid D)
=\frac{1}{BC}\sum_{c=1}^{C}\sum_{b=1}^{B}
\mathbf{1}\{\beta_{sj}^{(b,c)}\ne0\}.
\]

```{r selection}
lapply(inclusion_probabilities(fit), function(x)head(x, 3))
plot_top_features(fit, top = 5)
```

An overall ranking uses the largest subgroup probability for each feature;
it is not the probability of selection in at least one subgroup.

# Predict with the same fitted object

Prediction applies the training subgroup's centering and scaling. The following
uses training predictors to illustrate the return format; training predictions
do not measure performance on new subjects.

```{r predict}
prediction <- predict(fit, newdata = analysis, interval = TRUE)
lapply(prediction, head)
future <- predict(fit, newdata = analysis,
                  quantity = "new_observation", interval = TRUE, seed = 456)
lapply(future, head)
```

The default `quantity = "conditional_mean"` describes uncertainty in the
conditional response. `"new_observation"` also includes residual randomness
for a future subject. For binary models the default is a probability; future
outcomes are zero or one. For log-time survival, response-scale summaries are
medians and equal-tailed intervals of transformed draws. See `?predict.imr`
for the distinction between response and working scales.

# Use formulas for clinical covariates

The formula specifies the outcome and always-included clinical variables.
Molecular features remain in the platform tables. Formula preprocessing,
including factor levels and transformations, is saved for prediction.

```{r formula}
clinical <- merge(simIMR$outcome.continuous, simIMR$covariates, by = "id")
fit_age <- imr(y ~ age, data = clinical, platforms = simIMR$platforms,
  outcome_type = "continuous", priors = priors, mcmc = sampling)
fit_adjusted <- imr(y ~ age + sex + stage, data = clinical,
  platforms = simIMR$platforms, outcome_type = "continuous",
  priors = priors, mcmc = sampling)
compare_fit_summaries(list(age_only = fit_age, adjusted = fit_adjusted))
```

The comparison table describes fitted objects. Predictive comparison requires
held-out predictions. Use the same actual folds for both candidate formulas:

```{r paired-cv}
cv <- cv_imr(fit_age, k = 3, rounds = 1, cv_method = "refit", seed = 713)
cv_adjusted <- cv_imr(fit_adjusted, cv_method = "refit",
  folds = cv$control$folds, seed = 713)
cv
rbind(age_only = cv$pooled[, "all"], adjusted = cv_adjusted$pooled[, "all"])
head(cv$diagnostics)
```

Refitting repeats preprocessing and the complete joint MCMC within each training
fold. Fixed hyperparameters do not become data-independent if you chose them
using all outcomes. If CV itself selects the formula or hyperparameters, use
an outer evaluation layer as in the
[paired and nested comparison guide](covariate-comparison.html).

`cv_method = "reweight"` uses Pareto-smoothed importance sampling on stored
joint draws. It conditions on full-fit preprocessing and can be unreliable
when a held-out fold differs substantially from the full posterior. Its Pareto
k diagnostics must be checked; it is not an interchangeable faster refit.

# Binary and right-censored outcomes

The same interface and joint engine handle all three outcome types:

```{r other-outcomes}
binary_data <- imr_data(simIMR$platforms, simIMR$outcome.binary,
  simIMR$covariates, outcome_type = "binary")
binary_fit <- imr(binary_data, priors = priors, mcmc = sampling)
head(confint(binary_fit))

survival_data <- imr_data(simIMR$platforms, simIMR$outcome.survival,
  simIMR$covariates, outcome_type = "right.censored")
survival_fit <- imr(survival_data, priors = priors, mcmc = sampling,
  survival_scale = "log")
head(confint(survival_fit))
```

Binary sampling augments a Gaussian response constrained by the observed class.
Its variance prior anchors the latent scale near one. Survival sampling uses
observed log event times and augments right-censored log times above their
censoring bounds. These outcome models and the pMOM coefficient prior are
described in the [joint computation guide](joint-posterior.html).

# Extract samples and assess exploration

```{r draws}
beta <- posterior_draws(fit)
dim(beta[[1]])  # retained iteration, chain, coefficient
summary(fit, parm = "variance")
plot(fit, type = "coefficient_trace")
```

`posterior_draws()` only retrieves stored samples. It never starts additional
sampling. Multiple chains support rank-normalized split/folded R-hat and bulk
and tail effective sample sizes. Constant indicators are reported as undefined,
not as convergence successes. Examine traces, rare selection transitions and
Monte Carlo precision for the scientific quantities you intend to report.

# Reproduce an analysis

Keep the source snapshot, data, settings, saved chains, actual folds and seeds.
The [replay guide](reproducibility.html) shows what to record. Version 0.2.0
analyses used a different computational procedure; their saved selection-only
fits need refitting for this interface. The
[migration guide](https://h7nian.github.io/IntegMultiReg/migration.html)
lists the changes.

# References

Chekouo T, Stingo FC, Doecke JD, Do K-A (2017). A Bayesian integrative approach
for multi-platform genomic data: A kidney cancer case study.
*Biometrics*, **73**(2), 615–624.
[doi:10.1111/biom.12587](https://doi.org/10.1111/biom.12587).
