---
title: "Record and replay an analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Record and replay an analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A version label alone does not identify a development build. Install the source
snapshot recorded with an analysis, and retain the corresponding scripts and
reference outputs. The [migration guide](https://h7nian.github.io/IntegMultiReg/migration.html) maps interfaces used
in earlier snapshots to the current behavior-based names.

# Fix the source and environment

The source snapshot used by the current examples is
[`be38715b95568854e1a89140381ed4274f3102b2`](https://github.com/h7nian/IntegMultiReg/tree/be38715b95568854e1a89140381ed4274f3102b2). Install that immutable revision:

```{r install-snapshot, eval = FALSE}
remotes::install_github("h7nian/IntegMultiReg",
                        ref = "be38715b95568854e1a89140381ed4274f3102b2")
```

The locally built source archive used for the manuscript replay has SHA-256
`d86737a6dd51ef7589af941b100969bfaf26e984becd42f1016b0fa3f89400b7`.
This checksum identifies that exact built tarball, including generated vignettes;
a separately rebuilt archive or GitHub's automatic source download can have a
different checksum. Retain the original file when replaying the archived run.

Record the installed `RemoteSha` for a GitHub installation. For a source tarball,
keep the file and its SHA-256 checksum: `sha256sum IntegMultiReg_0.2.0.tar.gz` on
Linux, or `shasum -a 256 IntegMultiReg_0.2.0.tar.gz` on macOS. Also record R, GSL
and compiler versions. Fixed seeds do not guarantee identical floating-point
arithmetic across compilers and platforms.

```{r record-environment, eval = FALSE}
packageDescription("IntegMultiReg")$RemoteSha
sessionInfo()
```

# Keep data, transformations and actual partitions

The [getting started guide](IntegMultiReg.html) creates `analysis`, `fit` and
`cv` using the current symmetric-MRF updates and training-fold refits. Save the validated
inputs and full objects, not just rounded console output:

```{r retain-analysis, eval = FALSE}
saveRDS(list(data = analysis, fit = fit, cv = cv), "analysis.rds")
write.csv(cv$control$folds, "folds.csv", row.names = FALSE)
write.csv(cv$predictions, "out-of-fold-predictions.csv", row.names = FALSE)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```

The fit records its priors, preprocessing, sampler convention, chain lengths
and seed. CV controls record the actual folds and, for refit CV, the fitting
seeds in `refit_seeds`. For post-fit validation, retain `row_order` as well as
fold membership because floating-point summation order can affect the result.

Within a matching source and runtime, replay the recorded validation design:

```{r replay-folds, eval = FALSE}
saved <- readRDS("analysis.rds")
replay <- cv_imr(
  saved$fit, cv_method = "refit", folds = saved$cv$control$folds
)
stopifnot(identical(replay$predictions, saved$cv$predictions))
```

For a saved fit with retired convention names, explicitly call
`upgrade_imr_object()` before using it. The upgrade changes labels and validates
the structure; it preserves the original call and posterior draws. For exact
historical replay, retain the original source and scripts as well.

# Connect reference outputs to the manuscript

Keep a manifest linking each table or figure to its generating script, source
commit or source-package checksum, input-data checksum, numerical settings,
actual seeds/folds and unrounded output. Record which cells or figures the
manuscript and response letter use. Preserve previous references when an
algorithm changes, and explain the reason for a new numerical target before
accepting a replacement.

An independent repeat should run into a new directory without updating the
reference. Compare the saved predictions and scores, not only the printed
three-decimal table. The [covariate comparison guide](covariate-comparison.html)
shows how to reconstruct pooled MSE and check that outer test subjects never
enter inner validation.

# State what was verified

Successful execution and matching output establish computational repeatability
under the recorded conditions. They do not establish posterior convergence,
identify the cause of an older unexplained discrepancy, or confirm every
scientific conclusion of a study. Conditional coefficient-chain diagnostics
also do not diagnose the original selection chain. Keep the selection-chain
and ranking-stability limitations with the reported results.
