---
title: "Record and replay an analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Record and replay an analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A version label alone does not identify a development build. Retain the exact
source revision or built archive, data, scripts and reference outputs. Changing
the computational target requires a new numerical reference, even when the
statistical model is unchanged.

# Fix the source and environment

Record `packageDescription("IntegMultiReg")$RemoteSha` for a GitHub installation.
For a source tarball, keep the file and its SHA-256 checksum:
`sha256sum IntegMultiReg_0.3.0.tar.gz` on Linux or
`shasum -a 256 IntegMultiReg_0.3.0.tar.gz` on macOS. A rebuilt tarball can differ
in generated documentation or metadata, so keep the file actually installed.

```{r record-environment, eval=FALSE}
packageDescription("IntegMultiReg")$RemoteSha
sessionInfo()
RNGkind()
```

Also record the compiler and system. Fixed seeds do not guarantee identical
floating-point results across operating systems or compilers. Version 0.3.0
uses R's RNG and mathematical routines and no longer requires GSL.

# Keep inputs, chains and actual partitions

The [getting started guide](IntegMultiReg.html) constructs `analysis`, `fit_age`
and `cv`. Save full objects rather than rounded console output:

```{r retain-analysis, eval=FALSE}
saveRDS(list(data = analysis, fit = fit_age, cv = cv), "analysis.rds")
write.csv(cv$control$folds, "folds.csv", row.names = FALSE)
write.csv(cv$predictions, "out-of-fold-predictions.csv", row.names = FALSE)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```

The fit records resolved priors, subgroup transformations, chain lengths, RNG
kind, per-chain sampling and initialization seeds, and starting states. CV
records actual folds and, for refit CV, fold seeds and per-chain seeds. Workers
change execution scheduling, not the saved seed plan.

Within the same source and runtime, replay the recorded validation design:

```{r replay-folds, eval=FALSE}
saved <- readRDS("analysis.rds")
do.call(RNGkind, as.list(saved$cv$control$rng_kind))
replay <- cv_imr(saved$fit, cv_method = "refit",
  folds = saved$cv$control$folds, seed = saved$cv$control$seed)
stopifnot(identical(replay$predictions, saved$cv$predictions))
```

To replay a fit from its saved initialization, use the same data preparation,
formula, priors and RNG kind. Supply its recorded starts and root seed through
`imr_mcmc(initial = saved$fit$control$initial, seed = saved$fit$control$seed, ...)`,
with the original draws, burn-in, chain count and thinning. Initialization and
sampling have separate streams, so using the recorded starts does not shift
the sampling stream.

# Preserve earlier numerical targets

The previous 0.2.0 manuscript replay used source
[`be38715b95568854e1a89140381ed4274f3102b2`](https://github.com/h7nian/IntegMultiReg/tree/be38715b95568854e1a89140381ed4274f3102b2)
and the built archive SHA-256
`d86737a6dd51ef7589af941b100969bfaf26e984becd42f1016b0fa3f89400b7`.
That evidence belongs to the previous selection/conditional workflow. It does
not validate the 0.3.0 joint sampler or supply its reference outputs.

Selection-only saved objects cannot be converted into joint coefficient draws
by changing labels. Inspect historical fits with their archived package and
refit the original inputs to obtain a new joint posterior. The
[migration guide](https://h7nian.github.io/IntegMultiReg/migration.html)
describes the interface changes.

# Connect outputs to the manuscript

Keep a manifest linking each table or figure to its generating script, source
commit or archive checksum, data checksum, settings, actual seeds/folds and
unrounded output. Record which cells the manuscript and response letter use.
An independent repeat should write to a new directory without updating the
reference. Compare predictions and scores as well as displayed numbers.

Repeatability is distinct from statistical accuracy. Keep posterior exploration,
Monte Carlo precision and ranking-stability limitations with reported results.
Earlier conditional-chain diagnostics do not diagnose a joint chain. A small
new-to-old difference cannot by itself establish that an earlier discrepancy
was sampling fluctuation.
