---
title: "Record and replay an analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Record and replay an analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A version label alone does not identify a development build. The tutorials use
[source commit `641ca7a`](https://github.com/h7nian/IntegMultiReg/tree/641ca7a5bdb5b4004442732c5359ddeffa0e972e),
which names the sampler conventions `corrected` and `original` and uses
`postfit_original` for the original post-fit CV preset. The
[migration guide](../migration.html) maps values used in earlier snapshots.

# Fix the source and environment

```{r install-snapshot, eval = FALSE}
remotes::install_github(
  "h7nian/IntegMultiReg", ref = "641ca7a5bdb5b4004442732c5359ddeffa0e972e"
)
packageDescription("IntegMultiReg")$RemoteSha
sessionInfo()
```

For an analysis installed from a source tarball, retain that exact tarball and
its SHA-256 checksum alongside the generating scripts. On Linux use
`sha256sum IntegMultiReg_0.2.0.tar.gz`; on macOS use
`shasum -a 256 IntegMultiReg_0.2.0.tar.gz`. Also record the R, GSL and compiler
versions. A fixed seed does not guarantee identical floating-point arithmetic
across compilers and platforms.

# Keep data, transformations and actual partitions

The [getting started guide](IntegMultiReg.html) creates `analysis`, `fit` and
`cv` using the corrected sampler and training-fold refits. Save the validated
inputs and full objects, not just rounded console output:

```{r retain-analysis, eval = FALSE}
saveRDS(list(data = analysis, fit = fit, cv = cv), "analysis.rds")
write.csv(cv$control$folds, "folds.csv", row.names = FALSE)
write.csv(cv$predictions, "out-of-fold-predictions.csv", row.names = FALSE)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```

The fit records its priors, preprocessing, sampler convention, chain lengths
and seed. CV controls record the actual folds and, for refit CV, the fitting
seeds in `refit_seeds`. For post-fit validation, retain `row_order` as well as
fold membership because floating-point summation order can affect the result.

Within a matching source and runtime, replay the recorded validation design:

```{r replay-folds, eval = FALSE}
saved <- readRDS("analysis.rds")
replay <- cv_imr(
  saved$fit, cv_method = "refit", folds = saved$cv$control$folds
)
stopifnot(identical(replay$predictions, saved$cv$predictions))
```

For a saved fit with retired convention names, explicitly call
`upgrade_imr_fit()` before using it. The upgrade changes labels and validates
the structure; it preserves the original call and posterior draws. For exact
historical replay, retain the original source and scripts as well.

# Connect reference outputs to the manuscript

Keep a manifest linking each table or figure to its generating script, source
commit or source-package checksum, input-data checksum, numerical settings,
actual seeds/folds and unrounded output. Record which cells or figures the
manuscript and response letter use. Preserve previous references when an
algorithm changes, and explain the reason for a new numerical target before
accepting a replacement.

An independent repeat should run into a new directory without updating the
reference. Compare the saved predictions and scores, not only the printed
three-decimal table. The [covariate comparison guide](covariate-comparison.html)
shows how to reconstruct pooled MSE and check that outer test subjects never
enter inner validation.

# State what was verified

Successful execution and matching output establish computational repeatability
under the recorded conditions. They do not establish posterior convergence,
identify the cause of an older unexplained discrepancy, or confirm every
scientific conclusion of a study. Conditional coefficient-chain diagnostics
also do not diagnose the original selection chain. Keep the selection-chain
and ranking-stability limitations with the reported results.
